An Analysis of the Feasibility and Performance of BitTorrent as a Data Transfer Protocol in Peer to Peer Backup Systems

### 2014-04-30 paper/paper.tex
Paper: proofread chapters 4-7

\subsection{The Data Layer}
% TODO: find out how many unique secret prefixes there are, factoring that into the calculation of how many unique addresses there are.
To implement the data layer of our system, we use BitTorrent Sync. Each peer runs a copy of BitTorrent Sync as a daemon on their system. Our program interacts with it through the use of the BitTorrent Sync API \cite{btsyncapi}, several aspects of which we take advantage of. The first and main aspect is that is takes care of the actual data transfer; so long as our program provides BitTorrent Sync with the appropriate secrets and IP addresses of hosts involved in each file backup, it will take care of all data transfer in the data layer. The second is that the secrets that BitTorrent Sync generates can be used as general-purpose IDs in our system. We use these secrets as IDs for each peer when they join the overlay network, so that peers can be uniquely identified. This is possible due to the massive range of IDs that exist; according to the BitTorrent Sync Technology page, IDs must be at least 20 bytes long \cite{btsynctech}, which gives a minimum address space of $2^{160}$ unique addresses.

There are two downsides to using BitTorrent Sync. The first is that the software is currently closed source. This means that, while BitTorrent claims that BitTorrent Sync is completely secure, we have no know way of confirming this. It also means that we have less control over how our data layer functions and what information can be accessed from it. This is a problem that we had to solve with regard to the challenge mechanism described in Section \ref{subsec:ChallengeMechanism_sec:Fairness_chap:BTBackup}, the solution for which is described in Section \ref{subsubsec:MetadataController_subsec:SystemCore_sec:SystemDesign_chap:Implementation}. Ideally, we would find an open source version of BitTorrent Sync. However, we were not able to find any such systems that were fully functional. The second downside is that BitTorrent Sync is currently in Beta release. This means that, while the system is usable and has all of its features, there still might be many bugs, which leads to more error in our system as a whole.

@@

Through the use of the $backup$ command, the Console Controller is the part of our system that creates backups of files. In addition to finding nodes to create replicas on and asking them to back up the data, the file needs to be registered with BitTorrent Sync. There is a problem with this, however: data is registered in BitTorrent Sync by adding folders to it, not files. We want to treat each file as its own backed up data unit, with its own set of replica nodes. For this reason, each file that is backed up in our system needs to be in its own folder. We create a set of directories under a single root directory that is managed by our program. When a new file is added to BitTorrent Sync via our program, a new directory is created for it, and it is put into that folder.

@@

@@

\subsubsection{BTSync Controller} \label{subsubsec:BTSyncController_subsec:SystemCore_sec:SystemDesign_chap:Implementation}
% think about changing name to something more accurate
%Keeps the metadata layer up to date with local filesize changes of backed up data.
A challenge that is created by having our data layer implemented as a separate application is that the way the two programs view the network needs to be kept in sync. For example, when a file that is tracked by BitTorrent Sync is updated, the changes in content are automatically managed without the intervention of our program. Although BitTorrent Sync will see those changes and update its information accordingly, our program has no way of knowing that a change in file size has occurred. To remedy this, we created the BTSync Controller, whose purpose is to periodically query BitTorrent Sync through its API for changes in file size. Each peer stores a local copy of the metadata of its backed up files, which is compared to the data from the API query. If there is any difference in the file sizes, the controller concludes the file size has changed since the last check. In this case, it uses the peer to update the local copy of the metadata and send an update to the metadata layer to register the change in the network. Note that only file updates that change the file size cause the metadata layer to be updated, since our it is only concerned with the quantity of data being stored, rather than the content itself.

@@

In our implementation, the Metadata Controller periodically attempts to ping the nodes that are storing data for the peer in order to determine how reliable they are; this is a simplification of the challenge mechanism in Section \ref{subsec:ChallengeMechanism_sec:Fairness_chap:BTBackup}. When a node is deemed consistently unreliable, the Metadata Controller initiates the creation of a new backup and the deletion of the unreliable one.

@@

From these variables, we derived a set of test scenarios which allowed us to see how each of them affects the performance of the system. While they are specific to our system, they allow us to evaluate the strengths and weaknesses of BitTorrent Sync (and, by extension, BitTorrent) as the data transport mechanism in a peer-to-peer backup system.

@@

Each virtual machine was a headless GNU/Linux machine running Ubuntu 12.04, 64-bit. Each machine had a 2.1GB hard drive for the operating system, 1 virtual CPU, and 1GB of RAM. Depending on the test, we also attached another virtual hard drive to each virtual machine, as more storage was needed in order to backup large files.

Each machine is connected to each other via a virtual network that has a 1 Gbit link capacity. As this capacity is unrealistic for several of our tests, where the network is supposed to model users spread across the Internet, we used a bandwidth-limiting program called \textit{wondershaper}. \textit{wondershaper} is a command line utility that allows a user to set specific upload and download bandwidth limits on a per-interface basis. This is achieved by creating packet queues. When wondershaper is invoked to limit the bandwidth on a particular interface, it captures all of the packets going through that interface going in either direction. It can then remove them through the queue and send them in the appropriate direction, either towards the network or to a process on the same machine, at a rate specified by the user. The use of this utility allowed us to easily simulate realistic network delays for the tests that needed them.

@@

Our tests for this scenario used 8 virtual machines in SZTAKI Cloud: 1 peer, 5 nodes, 1 tracker, and 1 syslog-ng server. To have a wide range of file sizes, we tested with 10MB, 100MB, and 1GB files. In each test instance, the total amount of data backed up was 1GB, so 100 files, 10 files, and 1 file were backed up, respectively. The procedure for each test instance is as follows:

@@

At the end of this sequence, the peer has successfully used our system to backup its files and retrieve them when needed. On the peer and the nodes, the process was orchestrated by a script. Logging occurs during backup creation and retrieval. During both of these phases, the number of bytes being downloaded and uploaded was sampled once per second. This is the data that is sent to the log server. Sampling was achieved by using a series of shell commands to obtain the information and to put it in the useful format. The exact command is as follows:

@@

where \textit{eth0} is the name of the network interface. With the ability to query this at any time, we were able to determine the number of bytes sent each second by calling the above command once every second. The output is then sent to a logging server, which collects the data from every peer into one log file.

@@

In our test instance with churn, we modeled a network where each peer had an average uptime of 150 seconds (2.5 minutes) and an average downtime of 75 seconds (1.25 minutes). We chose these numbers based on the amount of time a typical user might be online in a given day. We reasoned that, if a typical user keeps his or her computer on during the day and shuts it off when he or she sleeps (for 8 hours), then the uptime in the average day will be 16 hours. Our implemented challenge mechanism sends a challenge message to replication nodes every 6 hours and blacklists them upon 50\% challenge failure over 4 days, meaning that a node is blacklisted if they fail 8 or more challenges. We scaled these numbers down so that we would have a 1 hour test: a 6 hour challenge becomes a 60 second challenge, and the challenge period becomes 16 minutes. To make the challenge period evenly fit into the 1 hour test, we rounded the challenge period down to 15 minutes. We scale 16 hours of uptime and 8 hours of downtime down to 2.5 minutes and 1.25 minutes, respectively.

@@

To make this test scenario more realistic, each node has their upload and download bandwidth limited to different values at the beginning of each instance. Download speeds are picked from a normal distribution of Internet speeds. In order to have a mean and standard deviation for our distribution that is reflective of current Internet speeds around the world, we found statistics on the average download speed per country, as well as the number of people in each country who use the Internet \cite{internetUsersByCountry,downloadSpeedsByCountry}. This allowed us to find the expected value for download speed by taking a weighted average, $W$:
$$
W=\sum\limits_{i=1}^{|\mathcal{D}|} n_iD_i
$$
Where $\mathcal{D}$ is the set of country download speeds and $n$ is the corresponding percentages of global Internet users that each country has. The normal distribution is only used for download speeds; rather than choosing an upload speed using the same method, we took each download speed and divided it by 2.527582524. This ensures that the upload speed is lower than the corresponding download speed, which is often the case on the Internet, as well as ensuring that the two speeds are proportional. We specifically used the aforementioned constant as that was the mean ratio between the two speeds based on the statistics we found.

@@

In addition to giving us insight into the effects file size and the number of files has on the speed of data backup, we used this test scenario to determine the extent that BitTorrent increases the data transfer rate of file backup and recovery.

@@

\begin{table}
\begin{adjustwidth}{-.5in}{-.5in}
\centering
    \begin{tabular}{| l | l | l | l |}
    \hline
    & One 1 GB File & Ten 100 MB files & Hundred 10 MB files \\ \hline
    Mean Speed (kB/s)& 8144.84 & 7874.69 & 5559.32 \\ \hline
    Standard Deviation of Speed (kB/s) & 2564.97 & 1899.86 & 1649.70 \\ \hline
    Median Speed (kB/s)& 9121.41 & 8379.84 & 5874.48 \\ \hline
    Max Speed (kB/s) & 11478.79 & 10948.98 & 8242.56 \\ \hline
    Duration (sec) & 162 & 165 & 231 \\ \hline
    \end{tabular}
    \caption{Statistics on Peer Data Recovery}
    \label{tab:PeerRecoverySpeed}
\end{adjustwidth}
\end{table}

It is worth noting that the increase in speed does not scale linearly with the size of the file. Although the size of the files increased by the same order of magnitude between the trials, the increase in recovery speed seems to have diminishing returns. Table \ref{tab:PeerRecoverySpeed} provides relevant statistics. The mean recovery speed increased by 41.65\% between the "Hundred 10 MB Files" trial and the "Ten 100 MB Files" trial. Compare this to an increase of 3.43\% between the "Ten 100 MB Files" trial and the "One 1 GB File" trial. Looking at it another way, the total time it took to recover the file between the former trials improves by 40\%, whereas the recovery time between the latter trials improves by only 4.76\%.

@@

There are a two interesting characteristics of this graph. The first is that data transfer does not start immediately; rather, it starts after the 10 second mark. This happens because the peer must index the file before sending parts of it to each node, as part of the BitTorrent Protocol. The second characteristic is that the data transfer rates for each node vary greatly over time. This can also be attributed to the BitTorrent Protocol: the peer requests file chunks from specific nodes, which corresponds to the peaks in data transfer. With this same reasoning, the peer might not use every node all of the time, which would correspond to the points in the graph where the upload speed drops for specific nodes.

@@

\begin{table}
\begin{center}
    \begin{tabular}{| l | l | l | l | l | l |}
    \hline
    & Node 1 & Node 2 & Node 3 & Node 4 & Node 5 \\ \hline
    Mean & 2303.07 & 1293.71 & 1702.47 & 1556.95 & 571.41 \\ \hline
    Standard Deviation & 1921.11 & 1714.13 & 1852.66 & 1796.30 & 1000.72 \\ \hline
    Median & 2024.57 & 477.37 & 1187.60 & 1028.09 & 171.38 \\ \hline
    Max & 7632.07 & 7423.9 & 6936.94 & 7237.46 & 5216.15 \\ \hline
    \end{tabular}
    \caption{Statistics on Node Upload Speed (in kB/s)}
    \label{tab:NodeRecoverySpeed}
\end{center}
\end{table}

@@

\begin{table}
\begin{adjustwidth}{-.5in}{-.5in}
\centering
    \begin{tabular}{| l | l | l | l | l | l | l |}
    \hline
    & Node 1 & Node 2 & Node 3 & Node 4 & Node 5 & Peer \\ \hline
    Mean & 1500.61 & 1094.83 & 1347.10 & 1240.98 & 656.69 & 2768.05\\ \hline
    Standard Deviation & 1815.66 & 1342.76 & 1645.63 & 1648.66 & 896.84 & 2512.48 \\ \hline
    Median & 757.55 & 585.35 & 757.56 & 480.7 & 190.36 & 2268.65 \\ \hline
    Max & 9141.97 & 7399.65 & 11195.35 & 8950.34 & 6252.98 & 9902.18 \\ \hline
    \end{tabular}
\end{adjustwidth}
    \caption{Upload speed statistics of each entity in a 1 GB transfer (in kB/s) \label{tab:UploadBackupSpeed}}
\end{table}

@@

This scenario was comprised of 100 peers each replicating a 100mb file 5 times, meaning there was theoretically a total of 50 GB of data in the network, ignoring the effects of churn. Table \ref{tab:ChurnBandwidth} summarizes the actual amount of data in the network, with and without churn.

\begin{table}
\begin{center}
    \begin{tabular}{| l | l |}
    \hline
    Without Churn & With Churn \\ \hline
    58.00 GB & 72.27 GB\\ \hline
    \end{tabular}
    \caption{Amount of data transferred in a 100 node network over 1 hour}
    \label{tab:ChurnBandwidth}
\end{center}
\end{table}

First, notice that the actual amount of data sent in the network without churn was 58 GB, 8 GB more than the calculated theoretical value. Distributed over 100 machines, this means each peer sent an additional 80 MB of data on average into the network throughout the duration of the test. Extra network traffic produced by our system could have contributed to this discrepancy, including communications with the tracker and each other. BitTorrent Sync could also be responsible, since there is overhead associated with both the metadata it maintains to stay synchronized, and the BitTorrent protocol it uses (e.g., requesting chunks of data from ). Communications between the script that instrumented the test and the log server may also have generated a nontrivial amount of traffic.

@@

It is important to note that this number is specific to the conditions that we have created for this test scenario. It is possible that a greater or lesser amount of churn would be present in a production environment; the amount that we have chosen is an approximation. The users of our system would be the ones to determine the availability of the average peer in the network. For example, if many users have servers, NAS devices, or other machines that are constantly online and are contributing resources to the network, the amount of churn will be lower, and therefore less replicas will need to be moved on average. The network could also be mainly made up of users who keep their computers off for most of the day, in which case the amount of traffic generated would be much higher. Since we have modeled our churn rate off of the average user, the results of our tests should be reflective of what would be seen in a production system.

@@

Section \ref{subsec:UploadSpeedofDataBackup} showed that our system lessens the Peer's load of backing up a file by distributing the task to the nodes as well. This has the side effect of speeding up the backup process, because nodes upload the data they have at a given time to each other, better utilizing their download bandwidth. Due to the recurring need for replica replacement, the benefits of speeding up the backup process will be experienced throughout the backup's lifetime. Our experiment showed that the speed increase could be upwards of 300\%. It is important to consider the number of replicas per file here, because the replica count that is maintained affects the speed and bandwidth usage of the backup and restore procedures. As the number of replicas per file increases, more traffic must be generated to create and maintain those replicas, but the download speed when recovering will be greatly improved. Of course, the maximum recovery speed is limited by the peer's download bandwidth.

Section \ref{sec:EffectsofChurn_chap:Results} explained that, in the presence of an average amount of churn, the average peer would need to make a new replica of any given file every 11 days. This time interval is derived by taking the average amount of time between the creation of replicas in our test instance and scaling that number based on the scaling factor that we used for the entire test instance.

@@

As consumer technology improves and higher resolution content can be generated and used by more people, the need for higher capacity backup systems increases. Most backup systems that have gained popularity have a centralized model, where backups are created on the storage servers of the entity that is offering the service. While this system model is practical for backing up a small amount of data, it is not a suitable option for users who need to backup a large amount of data, because these services often come with a data cap. Additional space is only obtainable by paying a fee.

@@

There are several challenges to creating a robust and reliability P2P backup system, however. Without the existence of a central entity that is solely dedicated to storing the files of others, it is possible that one's backups could be inaccessible at any given time, since regular users will often shut their computers off. Keeping track of who stores what data and making sure that everyone is contributing fairly are also major concerns. In addition to these problems, P2P backup systems share another problem with centralized systems: if users want to backup large amounts of data, the process of transferring these data is lengthy and inconvenient.

@@

In addition to addressing the problem of slow backup speeds in particular, BTBackup is useful as a general tool for P2P backup research. As P2P file backup is still a relatively new and dynamic research area, having a tool to prototype backup systems is valuable. Our program is open source, so it is easy for anyone to download our code and make improvements to any part of the system.

To test whether or not our system provided an increase in file transfer speeds, in addition to providing solutions for the other challenges of P2P backup systems, we implemented a prototype of our system and tested it on a virtualized, 100 node network using MTA SZTAKI's cloud computing system. We found that it was able to handle node churn with a reasonable amount of overhead in network traffic, requiring the replacement of a replica every 11 days. More importantly, however, we found that our system allowed for a \~300\% improvement in both file backup and restoration speeds over systems were a single uploader is used. This would not be possible if these data transfers were only between a user and some other entity, whether it is the service provider in a centralized backup system or a single peer in another P2P backup system.

@@

An example solution for (2): the communication between peers during the "Finding Replication Nodes" phase of backing up a file could be extended to require both parties digitally sign the information. That is, when a peer informs a node it will be backing up a given file of a given size, the DHT can store an additional entry containing their digital signatures, binding them to the exchange. This provides non-repudiation for the amount of data a peer is backing up and a node is storing, each of which a malicious user is incentivized to lie about. This solution does not deal with the blacklisters data, though a similar scheme for that is easy to imagine.

To reduce the size of the DHT with the addition of these digital signatures, a chaining mechanism could be put in place, similar to the one in Bitcoin's blockchain. Instead of appending digital signatures to the corresponding DHT column, new digital signatures could sign the old ones and then overwrite them. This preserves their effect, but removes the associated storage cost. More thought needs to be given to this to design an optimal mechanism (and to determine if chaining is even required).

>>>>

\subsection{The Data Layer}
% TODO: find out how many unique secret prefixes there are, factoring that into the calculation of how many unique addresses there are.
To implement the data layer of our system, we use BitTorrent Sync. Each peer runs a copy of BitTorrent Sync as a daemon on their system. Our program interacts with it through the use of the BitTorrent Sync API \cite{btsyncapi}, several aspects of which we take advantage of. The first and main aspect is that it takes care of the actual data transfer; so long as our program provides BitTorrent Sync with the appropriate secrets and IP addresses of hosts involved in each file backup, it will take care of all data transfer in the data layer. The second is that the secrets that BitTorrent Sync generates can be used as general-purpose IDs in our system. We use these secrets as IDs for each peer when they join the overlay network, so that peers can be uniquely identified. This is possible due to the massive range of IDs that exist; according to the BitTorrent Sync Technology page, IDs must be at least 20 bytes long \cite{btsynctech}, which gives a minimum address space of $2^{160}$ unique addresses.

There are two downsides to using BitTorrent Sync. The first is that the software is currently closed source. This means that, while BitTorrent claims that BitTorrent Sync is completely secure, we have no know way of confirming this. It also means that we have less control over how our data layer functions and what information can be accessed from it. This is a problem that we had to solve with regard to the challenge mechanism described in Section \ref{subsec:ChallengeMechanism_sec:Fairness_chap:BTBackup}, the solution for which is described in Section \ref{subsubsec:MetadataController_subsec:SystemCore_sec:SystemDesign_chap:Implementation}. Ideally, we would find an open source version of BitTorrent Sync. However, we were not able to find such a system that was fully functional. The second downside is that BitTorrent Sync is currently in Beta release. This means that, while the system is usable and has all of its features, there still might be many bugs, which leads to more error in our system as a whole.

@@

Through the use of the $backup$ command, the Console Controller is the part of our system that creates backups of files. In addition to finding nodes to create replicas on and asking them to back up the data, the file needs to be registered with BitTorrent Sync as part of the backup process. There is a problem with this, however: data is registered in BitTorrent Sync by adding folders to it, not files. We want to treat each file as its own backed up data unit, with its own set of replica nodes. For this reason, each file that is backed up in our system needs to be in its own folder. We create a set of directories under a single root directory that is managed by our program. When a new file is added to BitTorrent Sync via our program, a new directory is created for it, and it is put into that folder.

@@

% TODO when rm command is added, mention the required use of it when a user wants to delete a file so that our program knows to delete this hardlink

@@

\subsubsection{BTSync Controller} \label{subsubsec:BTSyncController_subsec:SystemCore_sec:SystemDesign_chap:Implementation}
% think about changing name to something more accurate
%Keeps the metadata layer up to date with local filesize changes of backed up data.
A challenge that is created by having our data layer implemented as a separate application is that the way the two programs view the network needs to be kept in sync. For example, when a file that is tracked by BitTorrent Sync is updated, the changes in content are automatically managed without the intervention of our program. Although BitTorrent Sync will see those changes and update its information accordingly, our program has no way of knowing that a change in file size has occurred. To remedy this, we created the BTSync Controller, whose purpose is to periodically query BitTorrent Sync through its API for changes in file size. Each peer stores a local copy of the metadata of its backed up files, which is compared to the data from the API query. If there is any difference in the file sizes, the controller concludes the file size has changed since the last check. In this case, it uses the Peer to update the local copy of the metadata and send an update to the metadata layer to register the change in the network. Note that only file updates that change the file size cause the metadata layer to be updated, since our application is only concerned with the quantity of data being stored, rather than the content itself.

@@

In our implementation, the Metadata Controller periodically attempts to ping the nodes that are storing data for the peer in order to determine how reliable they are; this is a simplification of the challenge mechanism described in Section \ref{subsec:ChallengeMechanism_sec:Fairness_chap:BTBackup}. When a node is deemed consistently unreliable, the Metadata Controller initiates the creation of a new replica and the deletion of the unreliable one.

@@

From these variables, we derived a set of test scenarios that allowed us to judge the performance of the system. While the test scenarios are specific to our system, they allow us to evaluate the strengths and weaknesses of BitTorrent Sync (and, by extension, BitTorrent) as the data transport mechanism in a peer-to-peer backup system.

@@

Each virtual machine was a headless GNU/Linux machine running Ubuntu 12.04, 64-bit. Each machine had a 2.1 GB hard drive for the operating system, 1 virtual CPU, and 1 GB of RAM. Depending on the test, we also attached another virtual hard drive to each virtual machine, as more storage was needed in order to backup large files.

Each machine is connected to each other via a virtual network that has a 1 Gbit link capacity. As this capacity is unrealistic for several of our tests, where the network is supposed to model users spread across the Internet, we used a bandwidth-limiting program called \textit{wondershaper}. \textit{wondershaper} is a command line utility that allows a user to set specific upload and download bandwidth limits on a per-interface basis. This is achieved by creating packet queues. When \textit{wondershaper} is invoked to limit the bandwidth on a particular interface, it captures all of the packets going through that interface in both directions. It can then remove them through the queue and send them in the appropriate direction, either towards the network or to a process on the same machine, at a rate specified by the user. The use of this utility allowed us to easily simulate realistic network delays for the tests that needed them.

@@

Our tests for this scenario used 8 virtual machines in SZTAKI Cloud: 1 peer, 5 nodes, 1 tracker, and 1 syslog-ng server. To have a wide range of file sizes, we tested with 10 MB, 100 MB, and 1 GB files. In each test instance, the total amount of data backed up was 1 GB, so 100 files, 10 files, and 1 file were backed up, respectively. The procedure for each test instance is as follows:

@@

At the end of this sequence, the peer has successfully used our system to backup its files and retrieve them when needed. On the peer and the nodes, the process was orchestrated by a script. Logging occurs during backup creation and retrieval. During both of these phases, the number of bytes being downloaded and uploaded was sampled once per second. This is the data that is sent to the log server. Sampling was achieved by using a series of shell commands to obtain the information and put it in a useful format. The exact command is as follows:

@@

\textit{eth0} is the name of the network interface. With the ability to query this at any time, we were able to determine the number of bytes sent each second by calling the above command once every second. The output is then sent to a logging server, which collects the data from every peer into one log file.

@@

% TODO discuss "our implemented challenge mechanism" in metadata controller of btbackup design chapter, and remind reader of it here
In our test instance with churn, we modeled a network where each peer had an average uptime of 150 seconds (2.5 minutes) and an average downtime of 75 seconds (1.25 minutes). We chose these numbers based on the amount of time a typical user might be online in a given day. We reasoned that, if a typical user keeps his or her computer on during the day and shuts it off when he or she sleeps (for 8 hours), then the uptime in the average day will be 16 hours. Our implemented challenge mechanism sends a challenge message to replication nodes every 6 hours and blacklists them upon 50\% challenge failure over 4 days, meaning that a node is blacklisted if they fail 8 or more challenges. We scaled these numbers down so that we would have a 1 hour test: a 6 hour challenge interval becomes a 60 second challenge interval, and the challenge period becomes 16 minutes. To make the challenge period evenly fit into the 1 hour test, we rounded the challenge period down to 15 minutes. We scale 16 hours of uptime and 8 hours of downtime down to 2.5 minutes and 1.25 minutes, respectively.

@@

To make this test scenario more realistic, each node has their upload and download bandwidth limited to different values at the beginning of each instance. Download speeds are picked from a normal distribution of Internet speeds. In order to have a mean and standard deviation for our distribution that is reflective of current Internet speeds around the world, we found statistics on the average download speed per country, as well as the number of people in each country who use the Internet \cite{internetUsersByCountry,downloadSpeedsByCountry}. This allowed us to find the expected value for download speed by taking a weighted average, $W$:
$$
W=\sum\limits_{i=1}^{|\mathcal{D}|} n_iD_i
$$
$\mathcal{D}$ is the set of country download speeds and $n$ is the corresponding percentages of global Internet users that each country has. The normal distribution is only used for download speeds; rather than choosing an upload speed using the same method, we took each download speed and divided it by 2.53, the mean ratio between download and upload bandwidths based on the aforementioned statistics. This ensures that the upload speed is lower than the corresponding download speed, which is often the case on the Internet, as well as ensuring that the two speeds are proportional.

@@

In addition to giving us insight into the effects that file size and the number of files has on the speed of data backup, we used this test scenario to determine the extent that BitTorrent increases the data transfer rate of file backup and recovery.

@@

\begin{table}
\begin{adjustwidth}{-.5in}{-.5in}
\centering
    \begin{tabular}{| l | l | l | l |}
    \hline
    & One 1 GB File & Ten 100 MB files & Hundred 10 MB files \\ \hline
    Mean Speed (kB/s)& 8144.84 & 7874.69 & 5559.32 \\ \hline
    Standard Deviation of Speed (kB/s) & 2564.97 & 1899.86 & 1649.70 \\ \hline
    Median Speed (kB/s)& 9121.41 & 8379.84 & 5874.48 \\ \hline
    Max Speed (kB/s) & 11478.79 & 10948.98 & 8242.56 \\ \hline
    Duration (sec) & 162 & 165 & 231 \\ \hline
    \end{tabular}
    \caption{Statistics on Peer Data Recovery.}
    \label{tab:PeerRecoverySpeed}
\end{adjustwidth}
\end{table}

It is worth noting that the increase in speed does not scale linearly with the size of the file. Although the size of the files increased by the same order of magnitude between the trials, the increase in recovery speed seems to have diminishing returns. Table \ref{tab:PeerRecoverySpeed} provides relevant statistics. The mean recovery speed increased by 41.65\% between the "Hundred 10 MB Files" trial and the "Ten 100 MB Files" trial. Compare this to an increase of 3.43\% between the "Ten 100 MB Files" trial and the "One 1 GB File" trial. Looking at it another way, the total time it took to recover the files between the former trials improves by 40\%, whereas the recovery time between the latter trials improves by only 4.76\%.

@@

There are two interesting characteristics of this graph. The first is that data transfer does not start immediately; rather, it starts after the 10 second mark. This happens because the peer must index the file before sending parts of it to each node, as part of the BitTorrent Protocol. The second characteristic is that the data transfer rates for each node vary greatly over time. This can also be attributed to the BitTorrent Protocol: the peer requests file chunks from specific nodes, which corresponds to the peaks in data transfer. With this same reasoning, the peer might not use every node all of the time, which would correspond to the points in the graph where the upload speed drops for specific nodes.

@@

\begin{table}
\begin{center}
    \begin{tabular}{| l | l | l | l | l | l |}
    \hline
    & Node 1 & Node 2 & Node 3 & Node 4 & Node 5 \\ \hline
    Mean & 2303.07 & 1293.71 & 1702.47 & 1556.95 & 571.41 \\ \hline
    Standard Deviation & 1921.11 & 1714.13 & 1852.66 & 1796.30 & 1000.72 \\ \hline
    Median & 2024.57 & 477.37 & 1187.60 & 1028.09 & 171.38 \\ \hline
    Max & 7632.07 & 7423.9 & 6936.94 & 7237.46 & 5216.15 \\ \hline
    \end{tabular}
    \caption{Statistics on Node Upload Speed (in kB/s).}
    \label{tab:NodeRecoverySpeed}
\end{center}
\end{table}

@@

\begin{table}
\begin{adjustwidth}{-.5in}{-.5in}
\centering
    \begin{tabular}{| l | l | l | l | l | l | l |}
    \hline
    & Node 1 & Node 2 & Node 3 & Node 4 & Node 5 & Peer \\ \hline
    Mean & 1500.61 & 1094.83 & 1347.10 & 1240.98 & 656.69 & 2768.05\\ \hline
    Standard Deviation & 1815.66 & 1342.76 & 1645.63 & 1648.66 & 896.84 & 2512.48 \\ \hline
    Median & 757.55 & 585.35 & 757.56 & 480.7 & 190.36 & 2268.65 \\ \hline
    Max & 9141.97 & 7399.65 & 11195.35 & 8950.34 & 6252.98 & 9902.18 \\ \hline
    \end{tabular}
\end{adjustwidth}
    \caption{Upload speed statistics of each entity in a 1 GB transfer (in kB/s). \label{tab:UploadBackupSpeed}}
\end{table}

@@

This scenario was comprised of 100 peers each replicating a 100 MB file 5 times, meaning there was theoretically a total of 50 GB of data in the network, ignoring the effects of churn. Table \ref{tab:ChurnBandwidth} summarizes the actual amount of data in the network, with and without churn.

\begin{table}
\begin{center}
    \begin{tabular}{| l | l |}
    \hline
    Without Churn & With Churn \\ \hline
    58.00 GB & 72.27 GB\\ \hline
    \end{tabular}
    \caption{Amount of data transferred in a 100 node network over 1 hour.}
    \label{tab:ChurnBandwidth}
\end{center}
\end{table}

First, notice that the actual amount of data sent in the network without churn was 58 GB, 8 GB more than the calculated theoretical value. Distributed over 100 machines, this means each peer sent an additional 81.92 MB of data on average into the network throughout the duration of the test. Extra network traffic produced by our system could have contributed to this discrepancy, including communications between the peers and the tracker, and between the peers themselves. BitTorrent Sync could also be responsible, since there is overhead associated with both the metadata it maintains to stay synchronized, and the BitTorrent protocol it uses (e.g., requesting chunks of data from peers). Communications between the script that instrumented the test and the log server may also have generated a nontrivial amount of traffic.

@@

It is important to note that these numbers are specific to the conditions that we have created for this test scenario. It is possible that a greater or lesser amount of churn would be present in a production environment; the amount that we have chosen is an approximation. The users of our system would be the ones to determine the availability of the average peer in the network. For example, if many users have servers, NAS devices, or other machines that are constantly online and are contributing resources to the network, the amount of churn will be lower, and therefore less replicas will need to be moved on average. The network could also be mainly made up of users who keep their computers off for most of the day, in which case the amount of traffic generated would be much higher. Since we have modeled our churn rate off of the average user, the results of our tests should be reflective of what would be seen in a production system.

@@

Section \ref{subsec:UploadSpeedofDataBackup} showed that our system lessens the peer's load of backing up a file by distributing the task to the nodes as well. This has the side effect of speeding up the backup process, because nodes upload the data they have at a given time to each other, better utilizing their download bandwidth. Due to the recurring need for replica replacement, the benefits of speeding up the backup process will be experienced throughout the backup's lifetime. Our experiment showed that the speed increase could be upwards of 300\%. It is important to consider the number of replicas per file here, because the replica count that is maintained affects the speed and bandwidth usage of the backup and restore procedures. As the number of replicas per file increases, more traffic must be generated to create and maintain those replicas, but the download speed when recovering will significantly improve. Of course, the maximum recovery speed is limited by the peer's download bandwidth.

Section \ref{sec:EffectsofChurn_chap:Results} explained that, in the presence of an average amount of churn, the average peer would need to make a new replica of any given file every 11 days. This time interval is derived by taking the average amount of time between the creation of replicas in our test instance and scaling that number based on the scaling factor that we used for the entire test instance. Furthermore, the overhead caused by churn is roughly 150\% of the file being backed up.

@@

As consumer technology improves and higher resolution content can be generated and used by more people, the need for higher capacity backup systems increases. Most backup systems that have gained popularity have a centralized model, where backups are created on the storage servers of the entity that is offering the service. While this system model is practical for backing up a small amount of data, it is not a suitable option for users who need to backup a large amount of data, because these services often come with a data cap; additional space is only obtainable by paying a fee.

@@

There are several challenges to creating a robust and reliable P2P backup system, however. Without the existence of a central entity that is solely dedicated to storing the files of others, it is possible that one's backups could be inaccessible at any given time, since regular users will often shut their computers off. Keeping track of who stores what data and making sure that everyone is contributing fairly are also major concerns. In addition to these problems, P2P backup systems share another problem with centralized systems: if users want to backup large amounts of data, the process of transferring the data is lengthy and inconvenient.

@@

In addition to addressing the problem of slow backup speeds in particular, BTBackup is useful as a general tool for P2P backup research. As P2P file backup is still a relatively new and dynamic research area, having a tool to prototype backup systems is valuable. Our program is open source, so it is easy for anyone to download our code and make improvements to any part of the system.\footnote{Our code is available at https://github.com/marcgreen/p2pbackup.}

To test whether or not our system provided an increase in file transfer speeds, in addition to providing solutions for the other challenges of P2P backup systems, we implemented a prototype of our system and tested it on a virtualized, 100 node network using MTA SZTAKI's cloud computing system. We found that it was able to handle node churn with a reasonable amount of overhead in network traffic, requiring the replacement of a replica every 11 days. More importantly, however, we found that our system allowed for a \~300\% improvement in both file backup and restoration speeds over systems where a single uploader is used. This would not be possible if these data transfers were only between a user and some other entity, whether it is the service provider in a centralized backup system or a single peer in another P2P backup system.

@@

An example solution for (2): the communication between peers during the "Finding Replication Nodes" phase of backing up a file could be extended to require both parties digitally sign the information. That is, when a peer informs a node that it will be backing up a given file of a given size, the DHT can store an additional entry containing their digital signatures, binding them to the exchange. This provides non-repudiation for the amount of data a peer is backing up and a node is storing, each of which a malicious user is incentivized to lie about. This solution does not deal with the blacklisters data, though a similar scheme for that is easy to imagine.

To reduce the size of the DHT with the addition of these digital signatures, a chaining mechanism could be put in place, similar to the one in Bitcoin's blockchain. Instead of appending digital signatures to the corresponding DHT column, the content that the new digital signatures sign could include the old digital signatures. The old ones could then be overwritten. This preserves their effect, but removes the associated storage cost. More thought needs to be given to this to design an optimal mechanism (and to determine if chaining is even required).
